Fast response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks.

### 2015-08-20 manuscript/manuscript_wtc3.Rmd
spelling fixes

# Introduction
Understanding and predict carbon (C) uptake in forest ecosystems is crucially important for assessing the impacts of environmental change. Specifically, this involves accurate upscaling of leaf level measurements to predict whole canopy function. Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling problematic. Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor] and photosynthesis (A) responds to this changing irradiance until until biochemical limitation occurs [@evans1995carbon]. Consequently, leaves modify structural and functional properties to efficiently intercept variable light and enhance photosynthetic capacity [@vogelman1996leaves; @valladares2004photosynthetic]. To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].

@@

\
These theoretical distributions of N and water regulate leaf physiological traits that constrain A differently for sun and shade leaves. The balance between the costs and benefits of these leaf traits affecting photosynthetic capacity create trade-offs impacting canopy C gain [@givnish1988adaptation]. During A, CO~2~ diffuses from the atmosphere through stomata (g~s~), intercellular air space and into the chloroplast for fixation by Rubisco (mesophyll conductance, g~m~) [@tazoe2011using]. Trade-offs exists that constrain each of these diffusion pathways, although these are yet to be explictily quantified for g~m~. For optimal leaf C gain these pathways should act to minimize the resistance to CO~2~ diffusion while also limiting the intrinsic energy, water and construction costs which diminish the C return from A. As these processes should not be considered independent of each other and must be integrated with hydraulic constraints [@griffiths2013you] predicting their behaviour within canopies remains difficult.

@@

## Stomatal conductance and leaf water use efficiency
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004). Increases in g~s~ were postively correlated to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a). Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a). With increased light availability, increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.

@@

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Table 2). Mesophyll conductance was nearly proportional to A within leaf types and light treatments (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b). No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types. Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figure `r get1fig`b and `r get2fig`b, respectively) and only a weak negative relationship with g~m~ and increasing leaf temperature was detected (Figure `r get1fig`b).

## Variation in intercellular and chloroplastic CO~2~ concentrations
More open stomata in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a). The drawdown of CO~2~ from intercellular spaces to the cholorplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature]. This drawdown was similar between sun and shade leaves at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c). This was the result of the proportional relationship between g~m~ and A across all leaves. The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).

@@

\
**Figure 7**.  Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.  The solid line represents the the theoretical line for C3 plants from [@evans1986carbon].

>>>>

# Introduction
Understanding and predict carbon (C) uptake in forest ecosystems is crucially important for assessing the impacts of environmental change. Specifically, this involves accurate upscaling of leaf level measurements to predict whole canopy function. Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling problematic. Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor] and photosynthesis (A) responds to this changing irradiance until biochemical limitation occurs [@evans1995carbon]. Consequently, leaves modify structural and functional properties to efficiently intercept variable light and enhance photosynthetic capacity [@vogelman1996leaves; @valladares2004photosynthetic]. To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].

@@

\
These theoretical distributions of N and water regulate leaf physiological traits that constrain A differently for sun and shade leaves. The balance between the costs and benefits of these leaf traits affecting photosynthetic capacity create trade-offs impacting canopy C gain [@givnish1988adaptation]. During A, CO~2~ diffuses from the atmosphere through stomata (g~s~), intercellular air space and into the chloroplast for fixation by Rubisco (mesophyll conductance, g~m~) [@tazoe2011using]. Trade-offs exists that constrain each of these diffusion pathways, although these are yet to be explicitly quantified for g~m~. For optimal leaf C gain these pathways should act to minimize the resistance to CO~2~ diffusion while also limiting the intrinsic energy, water and construction costs which diminish the C return from A. As these processes should not be considered independent of each other and must be integrated with hydraulic constraints [@griffiths2013you] predicting their behaviour within canopies remains difficult.

@@

## Stomatal conductance and leaf water use efficiency
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004). Increases in g~s~ were positively correlated to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a). Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a). With increased light availability, increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.

@@

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Table 2). Mesophyll conductance was nearly proportional to A within leaf types and light treatments (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b). No differences in g~s~ or g~m~ were detected with the warming treatment within leaf types. Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figure `r get1fig`b and `r get2fig`b, respectively) and only a weak negative relationship with g~m~ and increasing leaf temperature was detected (Figure `r get1fig`b).

## Variation in intercellular and chloroplastic CO~2~ concentrations
More open stomata in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a). The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature]. This drawdown was similar between sun and shade leaves at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c). This was the result of the proportional relationship between g~m~ and A across all leaves. The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).

@@

\
**Figure 7**.  Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.  The solid line represents the theoretical line for C3 plants from [@evans1986carbon].
